An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-31 VLDB2026/main.tex

	\subsection{Vector Libraries and Databases}
	Many existing vector libraries, such as Faiss, and databases, such as Milvus, VBASE, and PASE, are designed specifically for ANN search, but they also support attribute filtering and range queries. We also experimented with them. However, since these libraries and databases do not rely on a single algorithm, their internal implementations are complex, using multiple methods and have detailed third-party documentation, so we will not elaborate on them here.

@@

\textit{\textbf{Index size.}} As shown in Figure~\ref{fig:index_size_mb_comparison_query1}, PASE exhibits significantly larger index sizes across all supported datasets, mainly due to its reliance on a large amount of metadata. For medium-scale datasets, other methods show moderate variation. However, on large-scale dataset, Graph-based algorithms such as ACORN, FilteredDiskANN, and UNG tend to exhibit a sharp increase in index size, primarily due to the large number of candidate neighbors they maintain. Overall, IVF-based methods tend to produce more compact indexes.

@@

	\includegraphics[width=0.95\textwidth]{figures/exp/exp_3_1.pdf}
	\caption{
		 \centering
		Effect of Attribute Distribution on Query Performance \\
		(Strong: robust to attribute distribution shifts; Weak: sensitive to attribute distribution shifts.)}

>>>>

	\subsection{Vector Libraries and Databases}
The vector library and databases discussed in the following are not explicitly designed for hybrid query, but they support both attribute filtering and range filtering.

Faiss\footnote{\textcolor{blue}{https://github.com/facebookresearch/faiss}} \cite{Faiss} is an efficient similarity search library. It uses an ID selector to filter data before performing approximate nearest neighbor search. This supports hybrid search.In Faiss, we use the HQI batch optimization strategy in the IVF index. It groups queries with the same filter conditions. Each group only runs the filter once. It uses efficient matrix operations to compute vector similarity in batches.

PASE\footnote{\textcolor{blue}{https://github.com/alipay/PASE}} \cite{pase} is a vector index plugin for PostgreSQL. It uses a "search first, then filter" strategy. But because the candidate set size is fixed, it may not return enough top-k results.

VBASE\footnote{\textcolor{blue}{https://github.com/microsoft/MSVBASE}} \cite{vbase} is also a vector plugin for PostgreSQL. It applies filters during the HNSW index traversal. This ensures enough results.

Milvus\footnote{\textcolor{blue}{https://github.com/milvus-io/milvus}} \cite{milvus} is a vector database built for large-scale similarity search. It filters data first, then performs approximate nearest neighbor search. This supports hybrid search.

@@

\textit{\textbf{Index size.}}
\textcolor{orange}{
	As shown in Figure~\ref{fig:index_size_mb_comparison_query1}, PASE exhibits significantly larger index sizes across all supported datasets, mainly due to its reliance on a large amount of metadata. For regular-scale datasets, other methods show moderate variation. However, on large-scale dataset, Graph-based algorithms such as ACORN, FilteredDiskANN, and UNG tend to exhibit a sharp increase in index size, primarily due to the large number of candidate neighbors they maintain. Overall, IVF-based methods tend to produce more compact indexes.
}

@@

	\includegraphics[width=0.95\textwidth]{figures/exp/exp_3_1.pdf}
	\caption{
		Effect of Attribute Distribution on Query Performance}
%		(Strong is robust to attribute distribution shifts, Weak is sensitive to attribute distribution shifts.)}
